Cultural Transmission and Technological Transitions during the Late Paleolithic in Korea

### 2022-10-22 analysis/paper/paper.qmd
editing methods: materials

On the other end of our spectrum, we have socially connected groups that occupied places close to other groups or had regular contact with other groups. Individuals or groups learned the technology of stemmed points by copying a model from another individual or group. Relatively high degrees of social connectivity provide frequent opportunities for observing others and acquiring information. In this scenario, the model, or ideal, stemmed point is likely to have been a highly successful and frequently chosen one, and learners copy all information about the point design as a package. In this social context, indirect bias is the dominant influence on transmission of lithic technology. If indirect bias was prominent in the Korean Paleolithic, we expect that stemmed points should be more standardized, and attributes on the stemmed points will be more correlated, and assemblages from multiple sites will show less variance between and within them.

@@

# prepare some values to use in the text below
total_number_of_sp <- sum(stemmed_points_per_site$n)
total_number_of_sites <-nrow(stemmed_points_per_site)
```

@@

After the first discovery in Seokjangri site in the 1960s, around 300 stemmed points have been found in nearly 30 sites across Korea (@fig-map). Most sites contain only a few points but a few sites have many more, such as Suyanggae (n = 55), Jingeuneul (n = 99), and Yongsandong (n = 38) [@Kim_2017]. Among these stemmed points, we selected those that are unbroken from the tip to the stem. We excluded artifacts that were recorded as stemmed points but lack a stem. This resulted in a sample of `r total_number_of_sp` stemmed points from 28 assemblages unearthed from `r total_number_of_sites` sites spanning the period 40-17 ka (@tbl-sp-number). The images of the stemmed points were obtained from published excavation reports and by direct photography during our research on local museum collections. We defined multiple assemblages in a site where artifact-bearing deposits were separated by culturally sterile deposits, or where distinct artifact-bearing stratigraphic units could be identified by major differences in the texture, color, and composition of the sedimentary deposits.

We used previously developed chronologies that identify three phases in the Korean Late Paleolithic [@seong2015diversity; @Park_2013]. These chronologies were built on the presence of blades, stemmed point blanks, toolkit composition and radiocarbon ages. We arranged our 28 assemblages containing stemmed points in chronological order. For example, Sachang site has two assemblages that are fitted into phase 1 and 2. Applying the chronology is useful because some sites such as Yonghodong, Goryeri, Jungmal, and Mungyeong have no radiocarbon dates due to various research situations and limitations, chronological order help us estimate the time periods of those assemblages as well as understand the morphological change of stemmed points associated with the introduction of blade technology. Based on the existence of blade and micro blade and types of stemmed point blanks, we divided the assemblages into three chronological phases: 1) stemmed points made out of flakes and no blades in assemblages, 2) stemmed points made out of blade or flake and the existence of blade in assemblages, and 3) stemmed points made out of blade and the existence of micro blade in assemblages (@tbl-korean-chronology).

>>>>

On the other end of our spectrum, we have socially connected groups that occupied places close to other groups or had regular contact with other groups. Relatively high degrees of social connectivity provide frequent opportunities for observing others and acquiring information. Individuals or groups learned the technology of stemmed points by copying a model from another individual or group. In this scenario, the model, or ideal, stemmed point is likely to have been a highly successful and frequently chosen one, and learners copy all information about the point design as a package. In this social context, indirect bias is the dominant influence on transmission of lithic technology. If indirect bias was prominent in the Korean Paleolithic, we expect that stemmed points should be more standardized, and attributes on the stemmed points will be more correlated, and assemblages from multiple sites will show less variance between and within them.

@@

# prepare some values to use in the text below
total_number_of_sp <- sum(stemmed_points_per_site$n)
total_number_of_sites <-nrow(stemmed_points_per_site)
total_number_of_assemblages <- NULL # ADD CODE HERE
```

@@

After the first discovery of stemmed points at the Seokjangri site in the 1960s, around 300 have been found in nearly 30 sites across Korea (@fig-map). Most sites contain only a few points but a few sites have many more, such as Suyanggae (n = 55), Jingeuneul (n = 99), and Yongsandong (n = 38) [@Kim_2017]. Among these stemmed points, we selected those that are unbroken from the tip to the stem. We excluded artifacts that were recorded as stemmed points but lack a stem. This resulted in a sample of `r total_number_of_sp` stemmed points from 28 assemblages unearthed from `r total_number_of_sites` sites spanning the period 40-17 ka (@tbl-sp-number). The images of the stemmed points were obtained from published excavation reports and by direct photography during our research on local museum collections. We defined multiple assemblages in a site where artifact-bearing deposits were separated by culturally sterile deposits, or where distinct artifact-bearing stratigraphic units could be identified by major differences in the texture, color, and composition of the sedimentary deposits.

We used previously developed chronologies that identify three phases in the Korean Late Paleolithic [@seong2015diversity; @Park_2013]. These chronologies were built on radiocarbon ages, and for sites without any, the presence/absence of blades, stemmed point blanks, and toolkit composition . Based on the existence of blade and micro blade and types of stemmed point blanks, we divided the assemblages into three chronological phases: 1) stemmed points made out of flakes and no blades in assemblages, 2) stemmed points made out of blade or flake and the existence of blade in assemblages, and 3) stemmed points made out of blade and the existence of micro blade in assemblages (@tbl-korean-chronology). Applying this division is necessary because some sites such as Yonghodong, Goryeri, Jungmal, and Mungyeong have no radiocarbon dates due to various research limitations. Inferring ages for these sites by analogy to the technological sequences at sites with dates is important for making maximal use of the available archaeological data to increase our sample size. We arranged our `r total_number_of_assemblages` assemblages containing stemmed points in chronological order to facilitate observations of change over time.
